---
title: "Assignment 03 - Lego Sales"
author: "DylanMcKenna"
date: '2021-06-04'
slug: assignment-03-lego-sales
categories: []
tags: []
---

# This is Assignment 3 analyzing the lego_sales.csv file.




```{r echo=FALSE, message= FALSE, warning=FALSE}
library(tidyverse)
lego_sales <- read_csv("lego_sales.csv")


#drop columns that are not requird
lego_cleaned <- lego_sales %>%
  select(first_name:age, theme:us_price, quantity) %>%
  arrange(age, last_name)

```

## Determine the 3 Most Common Names

As shown in the chart below, the top 3 names are Jackson, Jacob, and Joseph.

```{r warning=FALSE}

common_names <- lego_cleaned %>%
  group_by(first_name) %>%
  count(sort = TRUE, name = "count_names") %>%
  arrange(desc(count_names)) %>%
  head(10)


ggplot(common_names, aes( x= first_name, y = count_names)) +
  geom_col(stat = "Count")

```

## Determine the Top 3 Themes

As shown below, the top 3 themes are 1) Star Wars, 2) Nexo Knights, and 3) Mixel and Gear.
```{r warning=FALSE}
common_themes <- lego_cleaned %>%
  group_by(theme) %>%
  count(sort = TRUE, name = "count_themes") %>%
  arrange(desc(count_themes)) %>%
  head(10)
  

ggplot(common_themes, aes( x= theme, y = count_themes)) +
  geom_col(stat = "Count")
```

## Determine the Most Common Subtheme Purchased of the Most Common Theme

The most common theme is Star Wars.

The most common subtheme in Star Wars is The Force Awakens.

```{r}

common_subtheme <- lego_cleaned %>%
  filter(theme == "Star Wars") %>%
  group_by(subtheme) %>%
  count(sort = TRUE, name = "count_subthemes") %>%
  arrange(desc(count_subthemes)) %>%
  head(10)

ggplot(common_subtheme, aes(x = count_subthemes, y = subtheme)) +
  geom_col()
```


## Create a New Variable of Age Categories

The probability of a customer's age group is presented below.

```{r}

lego_cleaned <- lego_cleaned %>%
  mutate( age_group = case_when( age <= 18 ~ "18 and under",
                                 between(age, 19, 25) ~ "19 - 25",
                                 between(age, 26, 35) ~ "26 - 35",
                                 between(age, 36, 50) ~ "36 - 50",
                                 age >= 51 ~ "51 and over")) %>%
select(age_group, age:subtheme, name:quantity, name)

lego_agegroup <- lego_cleaned  %>% 
  count( age_group) %>%
  mutate( probability = n / sum(n))

ggplot(lego_agegroup, aes(x = age_group, y = probability )) +
  geom_col()





```

## Determine the Age Group that has Purchased the Largest Number of Lego Sets

The age group 36 - 50 has purchased the most Lego sets.

```{r}
lego_sets  <- lego_cleaned %>%
  count(age_group, sum(quantity)) %>%
  mutate( sets_purchased = n)


ggplot(lego_sets, aes( x = age_group, y = sets_purchased)) + 
  geom_col()
```

## Determine the Age Group that has Spent the Most Money on Lego

The age group that has spent the most money is the 36 - 50 age bracket.

This age group spent $ 9,532.87.

```{r }


lego_spending <- lego_cleaned %>%
mutate(us_cost = us_price * quantity) %>%
  group_by(age_group) %>%
  summarise( us_cost = sum(us_cost)) 
    



ggplot(lego_spending, aes(x = age_group, y = us_cost)) +
  geom_col()


```

## Create a Question

How much does each age group spend on each theme?

It is clear that Star Wars is the number one seller. The age range from 26-50 appears to be spending the most money on most themes and is most noticeable in the Star Wars theme.


```{r}

lego_themespending <- lego_cleaned %>%
mutate(us_cost = us_price * quantity) %>%
  group_by(age_group, theme) %>%
  select(age_group, theme, us_cost)
  


ggplot(lego_themespending, aes(x = us_cost, y = reorder(theme, -us_cost, sum), fill = age_group)) + 
  geom_col()
```

## Add an element to the above plot

The change I made to the above plot was to the chart labels. I added better looking names to the chart.


```{r}
ggplot(lego_themespending, aes(x = us_cost, y = reorder(theme, -us_cost, sum), fill = age_group)) + 
  geom_col() +
  labs(title = "Spending on Themes by Age Group", fill = "Age Group" ) +
  xlab("Cost in US Dollars ($)") +
  ylab("Lego Theme") +
  theme_bw()
  

```

