---
title: Recoding Data
author: DylanMcKenna
date: '2021-06-01'
slug: recoding-data
categories: []
tags: []
---


## Import relig-income.xlsx

```{r}
library(readxl)
library(tidyverse)
rel_inc <- read_excel("relig-income.xlsx")
```


## Rename religion and n columns

```{r}
rel_inc %>%
  rename( 
    religion = `Religious tradition`, 
    n = `Sample Size` 
  )
```

## Pivot Longer
Short form is good for analysis. Long form is good for visualization.

Is it the other way around?
```{r}
rel_inc %>%
  rename(
    religion = `Religious tradition`,
    n = `Sample Size`
  ) %>%
  pivot_longer( 
    cols = -c(religion, n),   # all but religion and n 
    names_to = "income",  
    values_to = "proportion" 
  )
```


## Import and Transform rel_inc

```{r}
rel_inc %>%
  rename(
    religion = `Religious tradition`,
    n = `Sample Size`
  ) %>%
  pivot_longer(
    cols = -c(religion, n), 
    names_to = "income", 
    values_to = "proportion"
  ) %>%
  mutate(frequency = round(proportion * n))
```


## Save data
Repace existing or save to new data frame.

```{r}
rel_inc_long <- rel_inc %>% 
  rename(
    religion = `Religious tradition`,
    n = `Sample Size`
  ) %>%
  pivot_longer(
    cols = -c(religion, n), 
    names_to = "income", 
    values_to = "proportion"
  ) %>%
  mutate(frequency = round(proportion * n))
```


## Visualize using barplot

```{r}
ggplot(rel_inc_long, aes(y = religion, x = frequency)) +
  geom_col()
```


## Recode Religion

```{r}
# If error shows here, must run all chunks
#separate into many chunks

rel_inc_long <- rel_inc_long %>%
  mutate(religion = case_when(
    religion == "Evangelical Protestant" ~ "Ev. Protestant",
    religion == "Historically Black Protestant" ~ "Hist. Black Protestant",
    religion == 'Unaffiliated (religious "nones")' ~ "Unaffiliated",
    TRUE                                           ~ religion
  ))

# Create an update barplot

ggplot(rel_inc_long, aes(y = religion, x = frequency)) +
  geom_col()

rel_inc_long <- rel_inc_long %>%
  mutate(religion = fct_rev(religion))

ggplot(rel_inc_long, aes(y = religion, x = frequency)) +
  geom_col()
```

## Fill Barplot with Income

```{r}
ggplot(rel_inc_long, aes(y = religion, x = frequency, fill = income)) + 
  geom_col()
```

## Change theme of the plot

```{r}
ggplot(rel_inc_long, aes(y = religion, x = frequency, fill = income)) +
  geom_col(position = "fill") +
  scale_fill_viridis_d() +
  theme_minimal()
```

## Move Legend to the bottom 

```{r}
ggplot(rel_inc_long, aes(y = religion, x = frequency, fill = income)) +
  geom_col(position = "fill") +
  scale_fill_viridis_d() +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Adjust the legend
```{r}
ggplot(rel_inc_long, aes(y = religion, x = frequency, fill = income)) +
  geom_col(position = "fill") +
  scale_fill_viridis_d() +
  theme_minimal() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE))
```
## Fix the labels

```{r}
ggplot(rel_inc_long, aes(y = religion, x = frequency, fill = income)) +
  geom_col(position = "fill") +
  scale_fill_viridis_d() +
  theme_minimal() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE)) +
  labs(
    x = "Proportion", y = "", 
    title = "Income distribution by religious group", 
    subtitle = "Source: Pew Research Center, Religious Landscape Study", 
    fill = "Income" 
    )
```

