---
title: Assignment 05 - Exploratory Data Analysis - V2
author: DylanMcKenna
date: '2021-06-26'
slug: assignment-05-exploratory-data-analysis-v2
categories:
  - ggplot2
  - tidyverse
  - shiny
tags: []
---

# Assignmnet 5 - Exploratory Data Analysis

This is the submission for Assignment 5 - EDA.

## Dataset: City of Chicago - Crime - 2001 to Present

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(shiny)
library(lubridate)
library(gghighlight)
library(janitor)
library(gghighlight)
library(ggrepel)
```

The dataset used for this assignment was the City of Chicago Crime Statistics data. To reduce the file size, the data is reflective of the year 2019 and the year 2020.  This data is taken from the Chicago Police Departments publicly available data. 


```{r echo=FALSE, message=FALSE}
crime_post <- read.csv("crime_recent_V1.csv")

crime_post <- crime_post %>%
  filter(year == "2020") %>%
  mutate(date_day = as.Date(date_day, format = "%Y-%m-%d"),
         month = month(date_day, label = TRUE, abbr = TRUE))

#write.csv(crime_post, "crime_recent_V1.csv", row.names = FALSE)
```

### Q1 - What are the most common types of crimes?

```{r echo=FALSE}
crime_primary <- crime_post %>%
  count(primary_type) %>%
  mutate(primary_count = n) %>%
  select(-n)

ggplot(crime_primary, aes(x = primary_count, y = primary_type)) +
  geom_col(width = 0.5, position = "dodge") +
  theme_classic() +
  gghighlight(max(primary_count)) +
  geom_text(size = 3, aes(color = primary_type, 
                              label = primary_count),
                nudge_x = -2000,
                vjust = 1.5,
                                ) +
  labs(
        title = ("Frequency of Crime Reports for 2020"),
        x = "Number of Reports", y = "Primary Type of Crime",
        caption = "Data Source: City of Chicago - Crimes 2001 to Present",
        color = "Top 5 Types of Crime") 
```


### Q2 - What are the most frequent subtypes of the top 5 crime types?

```{r echo=FALSE}

crime_secondary <- crime_post %>%
  filter(primary_type %in% c("BATTERY", "THEFT", "ASSAULT", "CRIMINAL DAMAGE", "DECEPTIVE PRACTICE")) %>%
  count(primary_type, description) %>%
 mutate(description_count = n) %>%
#  select(-n) %>%
  arrange(desc(n)) #%>%
  

  
crime_secondary <- crime_secondary %>%
  slice(1:10)
  

ggplot(crime_secondary, aes(x = description_count, y = description, fill = primary_type, 
                            #label = description_count)
                            )) +
  #geom_text_repel() +
  geom_col(width = 0.5, position = position_dodge(1)) +
      theme_classic() +
     # geom_text_repel(size = 3,
      #                #hjust = -1,
                #nudge_x = 0,
                #vjust =  0,
                #                ) +
      labs(
        title = ("Frequency of Crime Reports for 2020"),
        x = "Number of Reports", y = "Secondary Type of Crime",
        caption = "Data Source: City of Chicago - Crimes 2001 to Present",
        fill = "Primary Type of Crime") 

```

## Q3 - When are more crimes reported during the year?

```{r echo=FALSE}

crime_year <- crime_post %>%
  count(date_day) %>% 
  mutate(count_day = n,
         month = month(date_day))
         #month = (month), 
         #mutate(month = month.abb[crime_year$month])) 

#crime_year <- crime_year %>%
  


ggplot(crime_year, aes(x= date_day, y = count_day) ) +
  geom_line(aes(color = month), show.legend = FALSE) +
  labs(
        title = ("Trend of Crime Reports for 2020"),
        x = "Date", y = "Frequency of Crime Reports",
        caption = "Data Source: City of Chicago - Crimes 2001 to Present"
        ) +
  theme_classic() +
  scale_colour_viridis_b()


```

## Q4 - Which month had the most crime in 2020?

```{r echo=FALSE}

crime_month <- crime_post %>%
  group_by(month) %>%
  arrange(month) %>%
  count(month) %>%
  mutate(month_count = n)
         

ggplot(crime_month, aes(x = month, y = month_count)
                            ) +
  #geom_text_repel() +
  geom_col(width = 0.5, position = position_dodge(1), aes(fill = month), show.legend = FALSE ) +
      theme_classic() +
  gghighlight(max(month_count)) +
     # geom_text_repel(size = 3,
      #                #hjust = -1,
                #nudge_x = 0,
                #vjust =  0,
                #                ) +
      labs(
        title = ("Frequency of Crime Reports per Month for 2020"),
        x = "Month", y = "Number of Crime Reports",
        caption = "Data Source: City of Chicago - Crimes 2001 to Present")
        #fill = "Primary Type of Crime") 



```


Daily Crime Rate

```{r echo=FALSE}

crime_rate <- crime_post %>%
  count(date_day, month) %>% 
  mutate(count_rate = n, rate = (count_rate/(sum(count_rate))) * 100)
        


ggplot(crime_rate, aes(x= date_day, y = rate, fill = month ))+
  geom_col() + #aes(fill = month), show.legend = TRUE) +
  gghighlight() +
  labs(
        title = ("Daily Crime Rate for 2020"),
        x = "Date", y = "Frequency of Crime Reports",
        caption = "Data Source: City of Chicago - Crimes 2001 to Present",
        fill = "Month"
        ) +
  theme_classic() +
  scale_fill_brewer(palette = "Paired")
  #scale_colour_viridis_b()

```


Which crimes have the highest arrest rate?

```{r}

crime_arrest <- crime_post %>%
  group_by(primary_type) %>%
  count(primary_type, arrest) %>%
  mutate(arrest_count = n,
         arrest = case_when(arrest == "1" ~ "Arrest Made",
                            arrest == "0" ~ "No Arrest")) %>%
  summarise(arrest_rate = (arrest_count/sum(arrest_count)*100),
            arrest = arrest)

head(crime_arrest)

ggplot(crime_arrest, aes(x= arrest_rate, y = primary_type, fill = arrest ))+
  geom_col(color = "black", position = position_stack(reverse = TRUE)) +
  gghighlight() +
  labs(
        title = ("Probability of Arrest by Crime Type"),
        x = "Percentage", y = "Primary Crime",
        caption = "Data Source: City of Chicago - Crimes 2001 to Present",
        fill = "Arrest Status"
        ) +
  theme_classic() + 
  scale_fill_brewer(palette = "Set1")
  #scale_colour_viridis_b()

```




Which wards are the best and worst to live in, based on crime rate?

```{r}


```

