---
title: Data Visualization using ggplot2
author: DylanMcKenna
date: '2021-05-18'
slug: data-visualization-using-ggplot2
categories: []
tags: []
---

## Palmer Penguins Data Analysis
We summarize the Palmer Penguins data using ggplot2 and tidyverse.

The use of facets and colour to present data.

```{r warning=FALSE, echo=FALSE, message=FALSE}
library(palmerpenguins)
library(tidyverse)
ggplot(penguins, 
       aes(x = bill_depth_mm, y = bill_length_mm,
       colour = species)) +
  geom_point() +
  labs(title = "Bill depth and length",
       subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
       x = "Bill depth (mm)", 
       y = "Bill length (mm)",
       colour = "Species",
       shape = "Island",
       caption = "Source: Palmer Station LTER / palmerpenguins package") +
  facet_grid( species ~ island ) +
  scale_colour_viridis_d()


```


Also, we can use facets without colour to present data with no legend.

```{r warning=FALSE}
library(palmerpenguins)
library(tidyverse)
ggplot(penguins, 
       aes(x = bill_depth_mm, y = bill_length_mm)) + 
  geom_point() +
  facet_grid( species ~ island )
```


We can use shapes to display data.
```{r warning = FALSE}
library(palmerpenguins)
library(tidyverse)
ggplot(penguins,
       aes(x = bill_depth_mm, 
           y = bill_length_mm,
           colour = species,
           shape = species)) +
  geom_point() +
  scale_colour_viridis_d()
```


We can use size to represent another variable. However, for overlapping data this can be difficult to understand. 
```{r warning=FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm, 
           y = bill_length_mm,
           colour = species,
           shape = species,
           size = body_mass_g)) +
  geom_point() +
  scale_colour_viridis_d()
```

We can include another variable using transparency. Also, to better represent overlapping data, we can use transparency (alpha) to see the data points.
```{r warning=FALSE}
ggplot(penguins,
       aes(x = bill_depth_mm, 
           y = bill_length_mm,
           colour = species,
           shape = species,
           size = body_mass_g,
           alpha = flipper_length_mm)) +
  geom_point() +
  scale_colour_viridis_d()
```

We can include a facet wrap.
```{r warning=FALSE}
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm)) + 
  geom_point() +
  facet_wrap(~ species)
```

