---
title: Final Project
author: DylanMcKenna
date: '2021-06-27'
slug: final-project
categories:
  - ggplot2
  - shiny
  - tidyverse
tags: []

output:
  html_document:
    code_folding: hide
   
---

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(shiny)
library(lubridate)
library(gghighlight)
library(janitor)
library(gghighlight)
library(ggrepel)
library(weathermetrics)
library(leaflet)
library(tidymodels)
library(knitr)

```


## Description of Dataset

The dataset used for this project is the Daily Weather in the U.S., 2017.  The data is collected by 1589 stations throughout the United States and Canada. The collected data consists of 15 variables; however, only 8 variables were used for this project. 

 
```{r echo=FALSE, warning=FALSE, message=FALSE, }
# read in data
weather <- read.csv("weather.csv")

# rename variables
weather <- weather %>%
  rename(min_temp_f = TMIN,
         max_temp_f = TMAX,
         avg_temp_f = TAVG,
         avg_day_wind_mph = AWND,
         max_wind_5_dir_deg = WDF5,
         max_wind_5_spd_mph = WSF5,
         snowfall_in = SNOW,
         snowdepth_in = SNWD,
         precip_in = PRCP)
  
station_count <- weather %>%
  group_by(station) %>%
  count()


```


```{r echo=FALSE, warning=FALSE, message=FALSE}
# convert fahrenheit to celsius for temperature variables
weather <- weather %>%
  mutate(min_temp_c = fahrenheit.to.celsius(min_temp_f, round = 2),
         max_temp_c = fahrenheit.to.celsius(max_temp_f, round = 2),
         avg_temp_c = fahrenheit.to.celsius(avg_temp_f, round = 2)) %>%
  select(-c(min_temp_f, max_temp_f, avg_temp_f)) 



# convert date to Date
weather <- weather %>%
  mutate(date = ymd(date))


```

```{r echo=FALSE, warning=FALSE, message=FALSE}
weather_group_temp_max <- weather %>%
  group_by(station) %>%
  count(station, max_temp_c, latitude, longitude) %>%
  summarise(max_temp_c = max(max_temp_c),
            latitude = max(latitude),
            longitude = max(longitude)) %>%
  na.omit() %>%
  arrange(desc(max_temp_c)) %>%
  #remove the outlier that is greather than 125 degrees. possible forest fire
  slice(-c(1))
```


### How does the temperature change with latitude? 

It is common sense to expect the temperature to decrease moving farther from the equator. The assumption is that there is an inverse relationship between latitude and temperature. As the latitude increases, the temperature will decrease.

### How does the maximum observed temperature change in relation to latitude?

The change in the maximum observed temperature related to the latitude fits the following model, where $\beta_0$ is 44.796 and $\beta_1$ is -0.304.  Statistically, both values are significant to a level of confidence greater than 99.99%.


$$\hat{y}_{i} = \beta_0 + \beta_1~x_{i}$$
In this case, the maximum temperature decreases at a rate of 0.30451 for every increase in degree of latitude.

```{r echo=FALSE}

# Fit a line to the date
line_maxtemp_latitude <- linear_reg() %>%
  set_engine("lm") %>%
  fit(max_temp_c ~latitude  , data = weather_group_temp_max) %>%
  tidy()

chart_maxtemp_latitude <- ggplot(data = weather_group_temp_max, 
            aes(x = latitude, y = max_temp_c)) +
  geom_point(color = "#3399FF", alpha = 0.5) +
  #gghighlight(max(max_temp_c))+
  labs(
    title = "Latitude and Maximum Temperature Trend",
    x = "Latitude",
    y = "Maximum Temperature Observed (Celsius)"
    ) +
  coord_cartesian(xlim = c(0, 75), ylim = c(0, 75)) #+
  #theme(plot.subtitle = element_text(colour = "#E48957", face = "bold", size = rel(1.5)))

chart_maxtemp_latitude <- chart_maxtemp_latitude +
  geom_smooth(formula = y ~ x, method = "lm", color = "#FF0000", fullrange = TRUE, se = FALSE) +
  theme_bw() 

chart_maxtemp_latitude
```


### How does average daily temperature change in relation to latitude?

Since we know the maximum temperature decreases with an increase in latitude, we assume the average temperature also decreases with an increase in latitude. In the chart below, the average temperature does not appear to change with a change in latitude. 

The change in the average observed temperature related to the latitude fits the following model, where $\beta_0$ is 20.813 and $\beta_1$ is -0.00041, basically zero.  Statistically, both values are significant to a level of confidence of 90%. 
 
For this chart, the data should be analyzed for any outliers that may be affecting the results.


$$\hat{y}_{i} = \beta_0 + \beta_1~x_{i}$$
In this case, the average temperature does not really change with a change in degree of latitude.


```{r echo=FALSE}
weather_group_temp_avg <- weather %>%
  group_by(station) %>%
  count(station, avg_temp_c, latitude) %>%
  summarise(avg_temp_c = max(avg_temp_c),
            latitude = max(latitude)) %>%
  na.omit() 
  

line_avgtemp_latitude <- linear_reg() %>%
  set_engine("lm") %>%
  fit(avg_temp_c ~latitude  , data = weather_group_temp_avg) %>%
  tidy()

chart_avgtemp_latitude <- ggplot(data = weather_group_temp_avg, 
            aes(x = latitude, y = avg_temp_c)) +
  geom_point(color = "#3399FF", alpha = 0.5) +
  geom_smooth(formula = y ~ x, method = "lm", color = "#FF0000", fullrange = TRUE, se = FALSE) +
  labs(
    title = "Latitude and Average Temperature Trend",
    x = "Latitude",
    y = "Average Temperature Observed (Celsius)"
    ) +
  coord_cartesian(xlim = c(0, 75), ylim = c(0, 75)) +
  theme_bw() 

chart_avgtemp_latitude
```


What about elevation?


### How does elevation affect the average temperature?

The chart below represents the relationship between average daily temperature and elevation. In the chart, it is noticeable that as the elevation increases, the average daily temperature decreases.

The change in the average daily temperature related to elevation fits the following model, where $\beta_0$ is 26.348 and $\beta_1$ is -0.00273.   Statistically, both values are significant to a level of confidence greater than 99.99%.


$$\hat{y}_{i} = \beta_0 + \beta_1~x_{i}$$
In this case, the average daily temperature decreases at a rate of 0.00273 for every foot of increase in elevation. 

This can be quite a change. For example, a 1000 foot increase in elevation would drop the average daily temperature by 2.73 degrees Celsius.

```{r echo=FALSE}
weather_group_temp_elevation <- weather %>%
  group_by(station) %>%
  count(station, avg_temp_c, elevation) %>%
  summarise(avg_temp_c = max(avg_temp_c),
            elevation = max(elevation)) %>%
  na.omit() 

line_avgtemp_elevation <- linear_reg() %>%
  set_engine("lm") %>%
  fit(avg_temp_c ~ elevation  , data = weather_group_temp_elevation) %>%
  tidy()

chart_avgtemp_elevation <- ggplot(data = weather_group_temp_elevation, 
            aes(x = elevation, y = avg_temp_c)) +
  geom_point(color = "#3399FF", alpha = 0.5) +
  geom_smooth(formula = y ~ x, method = "lm", color = "#FF0000", fullrange = TRUE, se = FALSE) +
  labs(
    title = "Elevation and Average Temperature Trend",
    x = "Elevation",
    y = "Average Temperature Observed (Celsius)"
    ) +
  coord_cartesian(xlim = c(0, 3500), ylim = c(0, 50)) +
  theme_bw() 

chart_avgtemp_elevation
```


### What is the relationship between elevation and precipitation?

From looking at the chart, many stations located at or near sea-level have a greater chance at observing a large precipitation event.

The chart below represents the relationship between maximum observed precipitation and elevation. In the chart, it is noticeable that as the elevation increases, the maximum observed precipitation decreases.

The change in the average daily temperature related to elevation fits the following model, where $\beta_0$ is 2.723 and $\beta_1$ is -0.000448. Statistically, both values are significant to a level of confidence greater than 99.99%.


$$\hat{y}_{i} = \beta_0 + \beta_1~x_{i}$$
In this case, the maximum observed precipitation decreases at a rate of 0.000448 inches for every foot of increase in elevation. 

This can be quite a change. For example, a 1000 foot increase in elevation would drop the maximum observed precipitation by 0.4 inches.


```{r echo=FALSE}
weather_group_precip_elevation <- weather %>%
  group_by(station) %>%
  count(station, precip_in, elevation) %>%
  summarise(precip_in = max(precip_in),
            elevation = max(elevation)) %>%
  na.omit() #%>%
  #arrange(desc(max_temp_c)) %>%
  #remove the outlier that is greather than 125 degrees. possible forest fire
  #slice(-c(1))

line_precip_elevation <- linear_reg() %>%
  set_engine("lm") %>%
  fit(precip_in ~ elevation  , data = weather_group_precip_elevation) %>%
  tidy()

chart_precip_elevation <- ggplot(data = weather_group_precip_elevation, 
            aes(x = elevation, y = precip_in)) +
  geom_point(color = "#3399FF", alpha = 0.5) +
  geom_smooth(formula = y ~ x, method = "lm", color = "#FF0000", fullrange = TRUE, se = FALSE) +
  labs(
    title = "Elevation and Maximum Precipitation Trend",
    x = "Elevation",
    y = "Maximum Observed Precipiation (Inches)"
    ) +
  coord_cartesian(xlim = c(0, 3500), ylim = c(0, 20)) +
  theme_bw() 

chart_precip_elevation
```

### Is there more precipitation with an increase in latitude?
According to the chart below, no, as latitude increases, the amount of precipitation decreases.

The change in the average daily temperature related to elevation fits the following model, where $\beta_0$ is 6.448 and $\beta_1$ is -0.1027. Statistically, both values are significant to a level of confidence greater than 99.99%.


$$\hat{y}_{i} = \beta_0 + \beta_1~x_{i}$$
In this case, the maximum observed precipitation decreases at a rate of 0.1027 inches for every increase in degree of latitude.


```{r echo=FALSE}
weather_group_precip_latitude <- weather %>%
  group_by(station) %>%
  count(station, precip_in, latitude) %>%
  summarise(precip_in = max(precip_in),
            latitude = max(latitude)) %>%
  na.omit() #%>%
  #arrange(desc(max_temp_c)) %>%
  #remove the outlier that is greather than 125 degrees. possible forest fire
  #slice(-c(1))

line_precip_latitude <- linear_reg() %>%
  set_engine("lm") %>%
  fit(precip_in ~ latitude  , data = weather_group_precip_latitude) %>%
  tidy()

chart_precip_latitude <- ggplot(data = weather_group_precip_latitude, 
            aes(x = latitude, y = precip_in)) +
  geom_point(color = "#3399FF", alpha = 0.5) +
  geom_smooth(formula = y ~ x, method = "lm", color = "#FF0000", fullrange = TRUE, se = FALSE, 
              ) +
  labs(
    title = "Latitude and Maximum Precipitation Trend",
    x = "Latitude",
    y = "Maximum Observed Precipiation (Inches)"
    ) +
  coord_cartesian(xlim = c(0, 75), ylim = c(0, 20)) +
  theme_bw() 

chart_precip_latitude
```


### Does maximum windspeed change with increasing latitude?

According to the chart below, yes, as latitude increases, the maximum windspeed decreases.

The change in the maximum windspeed related to elevation fits the following model, where $\beta_0$ is 64.2639 and $\beta_1$ is -0.1481. Statistically, both values are significant to a level of confidence greater than 90%.


$$\hat{y}_{i} = \beta_0 + \beta_1~x_{i}$$
In this case, the maximum observed windspeed decreases at a rate of 0.1481 inches for every increase in degree of latitude.

```{r echo=FALSE}
weather_group_wind_latitude <- weather %>%
  group_by(station) %>%
  count(station, max_wind_5_spd_mph, latitude) %>%
  summarise(max_wind_5_spd_mph = max(max_wind_5_spd_mph),
            latitude = max(latitude)) %>%
  na.omit() #%>%
  #arrange(desc(max_temp_c)) %>%
  #remove the outlier that is greather than 125 degrees. possible forest fire
  #slice(-c(1))

line_wind_latitude <- linear_reg() %>%
  set_engine("lm") %>%
  fit(max_wind_5_spd_mph ~ latitude  , data = weather_group_wind_latitude) %>%
  tidy()

chart_wind_latitude <- ggplot(data = weather_group_wind_latitude, 
            aes(x = latitude, y = max_wind_5_spd_mph)) +
  geom_point(color = "#3399FF", alpha = 0.5) +
  geom_smooth(formula = y ~ x, method = "lm", color = "#FF0000", fullrange = TRUE, se = FALSE, 
              ) +
  labs(
    title = "Latitude and Maximum 5-second Wind Speed Trend",
    x = "Latitude",
    y = "Maximum 5-second Wind Speed Observed (mph)"
    ) +
  coord_cartesian(xlim = c(0, 75), ylim = c(0, 150)) +
  theme_bw() 

chart_wind_latitude
```


