---
title: Final Project
author: DylanMcKenna
date: '2021-06-27'
slug: final-project
categories:
  - ggplot2
  - shiny
  - tidyverse
tags: []
output:
  html_document:
    code_folding: hide
---

```{r}
library(tidyverse)
library(ggplot2)
library(shiny)
library(lubridate)
library(gghighlight)
library(janitor)
library(gghighlight)
library(ggrepel)
library(weathermetrics)
library(leaflet)
library(tidymodels)

```


## Description of Dataset

The dataset used for this project is the Daily Weather in the U.S., 2017.  The data is collected by 1589 stations throughout the United States and Canada. The collected data consists of 15 variables; however, only 8 variables were used for this project. 
 
```{r warning=FALSE, error=FALSE}
# read in data
weather <- read.csv("weather.csv")

# rename variables
weather <- weather %>%
  rename(min_temp_f = TMIN,
         max_temp_f = TMAX,
         avg_temp_f = TAVG,
         avg_day_wind_mph = AWND,
         max_wind_5_dir_deg = WDF5,
         max_wind_5_spd_mph = WSF5,
         snowfall_in = SNOW,
         snowdepth_in = SNWD,
         precip_in = PRCP)
  
station_count <- weather %>%
  group_by(station) %>%
  count()

#str(station_count)

```


```{r echo=FALSE, warning=FALSE, message=FALSE}
# convert fahrenheit to celsius for temperature variables
weather <- weather %>%
  mutate(min_temp_c = fahrenheit.to.celsius(min_temp_f, round = 2),
         max_temp_c = fahrenheit.to.celsius(max_temp_f, round = 2),
         avg_temp_c = fahrenheit.to.celsius(avg_temp_f, round = 2)) %>%
  select(-c(min_temp_f, max_temp_f, avg_temp_f)) 



# convert date to Date
weather <- weather %>%
  mutate(date = ymd(date))


```

```{r echo=FALSE, warning=FALSE, message=FALSE}
weather_group_temp_max <- weather %>%
  group_by(station) %>%
  count(station, max_temp_c, latitude, longitude) %>%
  summarise(max_temp_c = max(max_temp_c),
            latitude = max(latitude),
            longitude = max(longitude)) %>%
  na.omit() %>%
  arrange(desc(max_temp_c)) %>%
  #remove the outlier that is greather than 125 degrees. possible forest fire
  slice(-c(1))
```


## How does the temperature change with latitude? 

Maximum temperature and latitude

```{r echo=FALSE}

# Fit a line to the date
line_maxtemp_latitude <- linear_reg() %>%
  set_engine("lm") %>%
  fit(max_temp_c ~latitude  , data = weather_group_temp_max) %>%
  tidy()

chart_maxtemp_latitude <- ggplot(data = weather_group_temp_max, 
            aes(x = latitude, y = max_temp_c)) +
  geom_point(color = "#3399FF", alpha = 0.5) +
  #gghighlight(max(max_temp_c))+
  labs(
    title = "Latitude and Maximum Temperature Trend",
    x = "Latitude",
    y = "Maximum Temperature Observed (Celsius)"
    ) +
  coord_cartesian(xlim = c(0, 75), ylim = c(0, 75)) #+
  #theme(plot.subtitle = element_text(colour = "#E48957", face = "bold", size = rel(1.5)))

chart_maxtemp_latitude <- chart_maxtemp_latitude +
  geom_smooth(formula = y ~ x, method = "lm", color = "#FF0000", fullrange = TRUE, se = FALSE) +
  theme_bw() 

chart_maxtemp_latitude
```


Average Temp by Latitude


```{r echo=FALSE}
weather_group_temp_avg <- weather %>%
  group_by(station) %>%
  count(station, avg_temp_c, latitude) %>%
  summarise(avg_temp_c = max(avg_temp_c),
            latitude = max(latitude)) %>%
  na.omit() #%>%
  #arrange(desc(max_temp_c)) %>%
  #remove the outlier that is greather than 125 degrees. possible forest fire
  #slice(-c(1))

line_avgtemp_latitude <- linear_reg() %>%
  set_engine("lm") %>%
  fit(avg_temp_c ~latitude  , data = weather_group_temp_avg) %>%
  tidy()

chart_avgtemp_latitude <- ggplot(data = weather_group_temp_avg, 
            aes(x = latitude, y = avg_temp_c)) +
  geom_point(color = "#3399FF", alpha = 0.5) +
  geom_smooth(formula = y ~ x, method = "lm", color = "#FF0000", fullrange = TRUE, se = FALSE) +
  labs(
    title = "Latitude and Average Temperature Trend",
    x = "Latitude",
    y = "Average Temperature Observed (Celsius)"
    ) +
  coord_cartesian(xlim = c(0, 75), ylim = c(0, 75)) +
  theme_bw() 

chart_avgtemp_latitude
```

Average Temp by Elevation

```{r echo=FALSE}
weather_group_temp_elevation <- weather %>%
  group_by(station) %>%
  count(station, avg_temp_c, elevation) %>%
  summarise(avg_temp_c = max(avg_temp_c),
            elevation = max(elevation)) %>%
  na.omit() #%>%
  #arrange(desc(max_temp_c)) %>%
  #remove the outlier that is greather than 125 degrees. possible forest fire
  #slice(-c(1))

line_avgtemp_elevation <- linear_reg() %>%
  set_engine("lm") %>%
  fit(avg_temp_c ~ elevation  , data = weather_group_temp_elevation) %>%
  tidy()

chart_avgtemp_elevation <- ggplot(data = weather_group_temp_elevation, 
            aes(x = elevation, y = avg_temp_c)) +
  geom_point(color = "#3399FF", alpha = 0.5) +
  geom_smooth(formula = y ~ x, method = "lm", color = "#FF0000", fullrange = TRUE, se = FALSE) +
  labs(
    title = "Elevation and Average Temperature Trend",
    x = "Elevation",
    y = "Average Temperature Observed (Celsius)"
    ) +
  coord_cartesian(xlim = c(0, 3500), ylim = c(0, 50)) +
  theme_bw() 

chart_avgtemp_elevation
```


Precipitation by Elevation

```{r echo=FALSE}
weather_group_precip_elevation <- weather %>%
  group_by(station) %>%
  count(station, precip_in, elevation) %>%
  summarise(precip_in = max(precip_in),
            elevation = max(elevation)) %>%
  na.omit() #%>%
  #arrange(desc(max_temp_c)) %>%
  #remove the outlier that is greather than 125 degrees. possible forest fire
  #slice(-c(1))

line_avgtemp_elevation <- linear_reg() %>%
  set_engine("lm") %>%
  fit(precip_in ~ elevation  , data = weather_group_precip_elevation) %>%
  tidy()

chart_precip_elevation <- ggplot(data = weather_group_precip_elevation, 
            aes(x = elevation, y = precip_in)) +
  geom_point(color = "#3399FF", alpha = 0.5) +
  geom_smooth(formula = y ~ x, method = "lm", color = "#FF0000", fullrange = TRUE, se = FALSE) +
  labs(
    title = "Elevation and Maximum Precipitation Trend",
    x = "Elevation",
    y = "Maximum Observed Precipiation (Inches)"
    ) +
  coord_cartesian(xlim = c(0, 3500), ylim = c(0, 20)) +
  theme_bw() 

chart_precip_elevation
```

Precipitation by Latitude

```{r echo=FALSE}
weather_group_precip_latitude <- weather %>%
  group_by(station) %>%
  count(station, precip_in, latitude) %>%
  summarise(precip_in = max(precip_in),
            latitude = max(latitude)) %>%
  na.omit() #%>%
  #arrange(desc(max_temp_c)) %>%
  #remove the outlier that is greather than 125 degrees. possible forest fire
  #slice(-c(1))

line_avgtemp_elevation <- linear_reg() %>%
  set_engine("lm") %>%
  fit(precip_in ~ latitude  , data = weather_group_precip_latitude) %>%
  tidy()

chart_precip_latitude <- ggplot(data = weather_group_precip_latitude, 
            aes(x = latitude, y = precip_in)) +
  geom_point(color = "#3399FF", alpha = 0.5) +
  geom_smooth(formula = y ~ x, method = "lm", color = "#FF0000", fullrange = TRUE, se = FALSE, 
              ) +
  labs(
    title = "Latitude and Maximum Precipitation Trend",
    x = "Latitude",
    y = "Maximum Observed Precipiation (Inches)"
    ) +
  coord_cartesian(xlim = c(0, 75), ylim = c(0, 20)) +
  theme_bw() 

chart_precip_latitude
```


Maximum 5 second Windspeed by Latitude

```{r echo=FALSE}
weather_group_wind_latitude <- weather %>%
  group_by(station) %>%
  count(station, max_wind_5_spd_mph, latitude) %>%
  summarise(max_wind_5_spd_mph = max(max_wind_5_spd_mph),
            latitude = max(latitude)) %>%
  na.omit() #%>%
  #arrange(desc(max_temp_c)) %>%
  #remove the outlier that is greather than 125 degrees. possible forest fire
  #slice(-c(1))

line_wind_latitude <- linear_reg() %>%
  set_engine("lm") %>%
  fit(max_wind_5_spd_mph ~ latitude  , data = weather_group_wind_latitude) %>%
  tidy()

chart_wind_latitude <- ggplot(data = weather_group_wind_latitude, 
            aes(x = latitude, y = max_wind_5_spd_mph)) +
  geom_point(color = "#3399FF", alpha = 0.5) +
  geom_smooth(formula = y ~ x, method = "lm", color = "#FF0000", fullrange = TRUE, se = FALSE, 
              ) +
  labs(
    title = "Latitude and Maximum 5-second Wind Speed Trend",
    x = "Latitude",
    y = "Maximum 5-second Wind Speed Observed (mph)"
    ) +
  coord_cartesian(xlim = c(0, 75), ylim = c(0, 150)) +
  theme_bw() 

chart_wind_latitude
```


