---
title: Assignment 01 - GGPLOT2
author: DylanMcKenna
date: '2021-05-21'
slug: assignment-01-ggplot2
categories: []
tags: []
---

# at least two level two headers ## and at least one bulleted list with at least two items * (2 points)
# at least two categorical data visualization. The code that created it must not be displayed. (2 points)
# at least two numerical data visualization. The code that created it must be displayed. (2 points) 
# a narrative discussing what you find interesting along with any issues you might have had creating this blog


#  This assignment was completed using the data set: mlb_players_18

This data contains the offensive statistics of Major League Baseball (MLB) players during the 2018 season.


```{r error=FALSE, include=FALSE}

data("mlb_players_18")
head(mlb_players_18)
library(openintro)
library(tidyverse)
str(mlb_players_18)
mlb_stats <- mlb_players_18
```


## Show number of homeruns by team and by position.


```{r}
# Show the homeruns per team


ggplot(mlb_stats, aes(y = team, x = HR)) +
      geom_bar(stat = "identity") +
  labs(
    x = "Home Runs",
    y = "MLB Team",
    title = "Home Runs by Each Team"
  )
  theme_classic()
         
```


```{r}
# Show the homeruns per position


ggplot(mlb_stats, aes(y = position, x = HR)) +
      geom_bar(stat = "identity") +
  labs(
    x = "Home Runs",
    y = "Position",
    title = "Home Runs by Each Position"
  )
  theme_classic()
```


```{r}
library(openintro)
pos_players <- subset(mlb_players_18, !position %in% c("P", "DH") & AB >= 100)
head(pos_players,10)

```



## Show homeruns by team, by position.


