---
title: Assignment 01 - GGPLOT2
author: DylanMcKenna
date: '2021-05-21'
slug: assignment-01-ggplot2
categories: []
tags: []
---


#  This assignment was completed using the data set: mlb_players_18

This data contains the offensive statistics of Major League Baseball (MLB) players during the 2018 season.


```{r error=FALSE, include=FALSE}
mlb_stats <- mlb_players_18
data("mlb_stats")
head(mlb_stats)
library(openintro)
library(tidyverse)
str(mlb_stats)

```


## Show number of homeruns by team and by position.


```{r echo=FALSE}

# Show the homeruns per team


ggplot(mlb_stats, aes(y = team, x = HR)) +
      geom_bar(stat = "identity") +
  labs(
    x = "Home Runs",
    y = "MLB Team",
    title = "Home Runs by Each Team") +
  theme_classic()
         
```



```{r echo=FALSE}
# Show the homeruns per position


ggplot(mlb_stats, aes(y = position, x = HR)) +
      geom_bar(stat = "identity") +
  labs(
    x = "Home Runs",
    y = "Position",
    title = "Home Runs by Each Position") +
  theme_classic()
```

## GREATER THAN 200 ABs: Show homeruns vs strike outs, strike outs vs walks, at-bats vs

The following charts present data for all position players and designated hitters that had 150 or more at-bats.


```{r}
# Create subset of position players and designated hitters with more than 150 at-bats
pos_players <- subset(mlb_stats, !position %in% c("P") & AB >= 150)

# Create Plot of Homerun vs Strike out
ggplot(pos_players, aes(x = strike_outs, y = HR)) +
      geom_point(alpha = 0.4) +
  labs(
    x = "Strikeouts",
    y = "Home Runs",
    title = "Risk and Reward of Swinging Big") +
  theme_classic()
```


 
```{r}

# Create Plot of Walks vs Strike out

ggplot(pos_players, aes(x = strike_outs, y = walks)) +
      geom_point(alpha = 0.4) +
  labs(
    x = "Strikeouts",
    y = "Walks",
    title = "Risk and Reward of Seeing More Pitches") +
  theme_classic()
```


```{r}

# Show hits related to RBIs

ggplot(pos_players, aes(x = H, y = RBI)) +
      geom_point(alpha = 0.4) +
    labs(
    x = "Hits",
    y = "RBI",
    title = "Drive In Those Runs") + 
  theme_classic()
```


```{r}

# Show Runs related to Stolen Bases

ggplot(pos_players, aes(x = stolen_bases, y = R)) +
      geom_point(alpha = 0.4) +
    labs(
    x = "Stolen Bases",
    y = "Runs",
    title = "Risk and Reward of Stealing Bases") + 
  theme_classic()
```

