---
title: Assignment 01 - GGPLOT2
author: DylanMcKenna
date: '2021-05-21'
slug: assignment-01-ggplot2
categories: []
tags: []
---


#  This assignment was completed using the data set: mlb_players_18

This data contains the offensive statistics of Major League Baseball (MLB) players during the 2018 season.


```{r error=FALSE, echo=FALSE}


data("mlb_players_18")
head(mlb_players_18)
library(openintro)
library(tidyverse)
str(mlb_players_18)

```


## This section contain the categorical data.

Show number of homeruns by team and by position.


```{r echo=FALSE}
library(openintro)
library(tidyverse)
mlb_stats <- mlb_players_18

# Show the homeruns per team


ggplot(mlb_stats, aes(y = team, x = HR)) +
      geom_bar(stat = "identity") +
  labs(
    x = "Home Runs",
    y = "MLB Team",
    title = "Home Runs by Each Team") +
  theme_classic()
         
```


Show number of homeruns by position.


```{r echo=FALSE}

library(openintro)
library(tidyverse)

# Show the homeruns per position


ggplot(mlb_stats, aes(y = position, x = HR)) +
      geom_bar(stat = "identity") +
  labs(
    x = "Home Runs",
    y = "Position",
    title = "Home Runs by Each Position") +
  theme_classic()

```


## This section contains the numerical data.

The following charts present data for all position players and designated hitters that had 150 or more at-bats.

A player that had at least 150 at-bats would have played in enough games to get consistent playing time.

Home runs have an important impact on a game. Although, some players that swing for home runs can run a risk of striking out more frequently.

```{r}

# Create subset of position players and designated hitters with more than 150 at-bats

pos_players <- subset(mlb_stats, !position %in% c("P") & AB >= 150)

# Create Plot of Homerun vs Strike out

ggplot(pos_players, aes(x = strike_outs, y = HR)) +
      geom_point(alpha = 0.4) +
  labs(
    x = "Strikeouts",
    y = "Home Runs",
    title = "Risk and Reward of Swinging Big") +
  theme_classic()

```

Some players' strategy is to see more pitches. By seeing more pitches, the batter has an increased opportunity to earn a walk with the risk of striking out.

 
```{r}

# Create Plot of Walks vs Strike out

ggplot(pos_players, aes(x = strike_outs, y = walks)) +
      geom_point(alpha = 0.4) +
  labs(
    x = "Strikeouts",
    y = "Walks",
    title = "Risk and Reward of Seeing More Pitches") +
  theme_classic()

```

Do more hits increase the potential for RBI's?

Looking at the following chart, it appears that the more hits a player earns, the more RBI's a player earns.


```{r}

# Show hits related to RBIs

ggplot(pos_players, aes(x = H, y = RBI)) +
      geom_point(alpha = 0.4) +
    labs(
    x = "Hits",
    y = "RBI",
    title = "Drive In Those Runs") + 
  theme_classic()

```

Another area to explore is the impact of stealing bases on scoring more runs. 


```{r}

# Show Runs related to Stolen Bases

ggplot(pos_players, aes(x = stolen_bases, y = R)) +
      geom_point(alpha = 0.4) +
    labs(
    x = "Stolen Bases",
    y = "Runs",
    title = "Risk and Reward of Stealing Bases") + 
  theme_classic()

```


An issue that I have found creating this blog is figuring out why sometimes the information uploads to netlify and sometimes it does not.  

I find it interesting how we can present data with so few inputs.
