---
title: Visualising Numerical Data
author: Dylan McKenna
date: '2021-05-21'
slug: visualising-numerical
categories:
  - ggplot2
tags: []
---

# Lending Club Data Visualisation
We visualize the Lending Club data using ggplot2.

A glimpse at Lending Club data.

```{r warning=FALSE}
library(openintro)
glimpse(loans_full_schema)

```

```{r error=FALSE}
loans <- loans_full_schema %>%
  select(loan_amount, interest_rate, term, grade, 
         state, annual_income, homeownership, debt_to_income)
glimpse(loans)
```

## Histograms
We can check the number of loans for various amounts.

```{r}
ggplot(loans, aes(x = loan_amount)) +
  geom_histogram(binwidth = 1000)
```
We can try a larger bin width if desired.

```{r}
ggplot(loans, aes(x = loan_amount)) +
  geom_histogram(binwidth = 5000)
```

We can customize our histogram.
```{r}
ggplot(loans, aes(x = loan_amount)) +
  geom_histogram(binwidth = 5000) +
  labs(
    x = "Loan amount ($)",
    y = "Frequency",
    title = "Amounts of Lending Club loans"
  )
```

We can fill using a categorical variable.
```{r}
ggplot(loans, aes(x = loan_amount, 
                  fill = homeownership)) +
  geom_histogram(binwidth = 5000,
                 alpha = 0.5) +
  labs(
    x = "Loan amount ($)",
    y = "Frequency",
    title = "Amounts of Lending Club loans"
  )
```

We can facet with a categorical variable.
```{r}
ggplot(loans, aes(x = loan_amount, fill = homeownership)) + 
  geom_histogram(binwidth = 5000) +
  labs(
    x = "Loan amount ($)",
    y = "Frequency",
    title = "Amounts of Lending Club loans"
  ) +
  facet_wrap(~ homeownership, nrow = 3)
```

## Density Plots

We could use a density plot.
```{r}
ggplot(loans, aes(x = loan_amount)) +
  geom_density()

```
For density plots, the bandwidth can be adjusted.

The bandwith here is 0.5.
```{r}
ggplot(loans, aes(x = loan_amount)) +
  geom_density(adjust = 0.5)
```
The bandwidth here is 2.
```{r}
ggplot(loans, aes(x = loan_amount)) +
  geom_density(adjust = 2)
```
We can also customize the density plots.
```{r}
ggplot(loans, aes(x = loan_amount)) +
  geom_density(adjust = 2) +
  labs(
    x = "Loan amount ($)",
    y = "Density",
    title = "Amounts of Lending Club loans"
  )
```
We can add a categorical variable to our density plot.
```{r}
ggplot(loans, aes(x = loan_amount, 
                  fill = homeownership)) +
  geom_density(adjust = 2, 
               alpha = 0.5) +
  labs(
    x = "Loan amount ($)",
    y = "Density",
    title = "Amounts of Lending Club loans", 
    fill = "Homeownership"
  )
```

## Box Plots
We can also use box plots.

```{r}
ggplot(loans, aes(x = interest_rate)) +
  geom_boxplot()
```

Box plots can be used to show obvious outliers.
```{r}
ggplot(loans, aes(x = annual_income)) +
  geom_boxplot()
```

We can customize our box plots.

```{r}
ggplot(loans, aes(x = interest_rate)) +
  geom_boxplot() +
  labs(
    x = "Interest rate (%)",
    y = NULL,
    title = "Interest rates of Lending Club loans"
  ) +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  )
```

We can add a categorical variable.

```{r}
ggplot(loans, aes(x = interest_rate,
                  y = grade)) +
  geom_boxplot() +
  labs(
    x = "Interest rate (%)",
    y = "Grade",
    title = "Interest rates of Lending Club loans",
    subtitle = "by grade of loan"
  )
```

## Scatterplot

We can use a basic scatterplot.

```{r warning=FALSE}
ggplot(loans, aes(x = debt_to_income, y = interest_rate)) +
  geom_point()
```

We can change the marker shape. In the following example, the dots become hexagons.

```{r}
ggplot(loans %>% filter(debt_to_income < 100), 
       aes(x = debt_to_income, y = interest_rate)) +
  geom_hex()
```

