---
title: 'Assignment A02: TIDYVERSE'
author: DylanMcKenna
date: '2021-05-28'
slug: assignment-a02-tidyverse
categories:
  - tidyverse
tags: []
---


## Assignment 02: TIDYVERSE



## Timely Submission (2 points)


```{r echo=FALSE, message=FALSE, error=FALSE, warning= FALSE}
library(tidyverse)
library(dplyr)
library(convertr)

master <- read_csv("Master.csv")

batting <- read_csv("Batting.csv")

fielding <- read_csv("Fielding.csv")

  
```

```{r}

batting <- batting %>%
  mutate( GIDP = as.double(GIDP))

head(batting)

fielding <- fielding %>%
  mutate( lgID = as.character(lgID))

head(master)
tail(fielding)

```


## at least two functions each: one table verb and two table verb  (2 points)

Combine master into batting and into fielding.

```{r}
# Join master list into the batting list to show the players name.
batting_full <- left_join(batting, master)
player_data <- batting_full %>%
  full_join(fielding, by = "playerID")

tail(batting_full)
tail(player_data)
tail(player_data, 100)
```


## at least two functions each of Grouping and Vector functions (2 points)

## at least two functions each of Pivoting and Missing Values (tidyr). (2 points)

## at narrative discussing what you find interesting along with any issues you might have had creating this blog (2 points)
